|

### 2023-02-01 manuscript/manuscript_1.0.Rmd
Update the figures. Changing the order of the panels and adding labels

---
title: |
  | Learning can mediate both badges of status and individual recognition
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
- C. Daniel Cadena*
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  word_document: default
fig_caption: yes
fontsize: 12pt
header-includes: |
  ```{=latex}
    \usepackage{float}
    \usepackage{setspace}\doublespacing
    \usepackage[backref=page]{hyperref}
    \linespread{2}
    \usepackage{lineno}
    \linenumbers
    \floatplacement{figure}{H}
   \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  ```
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: ../SocialMorphCue.bib
spacing: double
style: C:\Users\andre\Zotero\styles\elife.csl
abstract: | 
  Adaptive behavioural responses often depend on qualities of the interacting partner of an individual. For example, when competing for resources, an individual might be better off escalating fights with individuals of lower quality, while restraining from fighting individuals of higher quality. Communication systems involving signals of quality allow individuals to reduce uncertainty regarding the fighting ability of their partners and make more adaptive behavioral decisions. However, dishonest individuals can destabilize such communications systems. One open question is whether cognitive mechanisms, such as learning, can maintain the honesty of signals, thus, favoring their evolutionary stability. Here we present evolutionary simulations, where individuals can produce a signal proportional to their quality; and learn along their lifetime the best response to the signal emitted by their peers. In these simulations, learning on the receiver side can mediate the evolution of signals of quality on the sender side. When the cost of the signal is proportional to the quality of the sender populations are only composed of honest signalers. When the cost is not proportional to the quality of the signaler, the population is composed of both honest and dishonest signalers. We argue that learning can be a general cognitive mechanism playing a role in a wide range of communication systems.  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,warnings=FALSE)
require(png)
require(here)
source(here("../R_files/posPlots.r"))
source(here("AccFunc.r"))
```

@@

The decision of whether to escalate a fight against an interacting 
partner can be dependent on the size of the partner's badge, 
the actual dependency is determined by the focal´s experiences 
through a learning process. We implement learning using 
the actor-critic approach from reinforcement learning theory [RL; @sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
Individuals estimate the reward (pay-off) expected from  interacting 
with partners of different badge sizes (the critic in RL terminology). 
After every interaction they update the estimate of reward proportionally 
to the difference between their current estimate and the observed 
reward (prediction error $\delta$) and to the speed of learning 
($\mathrm{A}$). Furthermore, they express different probabilities of 
retreating (attacking) depending on the badge size of their opponent 
(the actor in RL). They update the probability of 
retreating (attacking)  up or down depending on whether retreating 
(attacking) leads to an increase in the reward estimation. 
So, if a focal individual decides to escalate a fight against 
an individual with small badge, and that leads to an increase in 
the reward estimation, the focal individual will increase the probability 
of escalating fights with individuals of small badges in the future. 
Given that badge size is a real number between 0 and 1, there are infinitely 
many badge sizes. Thus, the reward estimation, as well as the probability 
of retreating (escalating), must be generalized across different values. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018]. 
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]). Each one of these features centres 
is associated with a weight for reward estimation and tendency to play 
retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights 
associated with each feature, weighted by the response 
triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
represent the feature weights) The response of each feature centre 
diminishes as a Gaussian function  with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:model-struc} B, grey line). Formally the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,
\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
\end{equation}

@@

\begin{equation}
logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
\end{equation}

@@

After all their interactions, individuals in the population 
reproduce proportionally to their fitness $w_i$, which is a sum of the 
baseline fitness ($w_0$) and all the pay-offs obtained throughout their 
life. Thus, the combination of natural selection and genetic drift 
changes the distribution of values in $\alpha$ and $\beta$ that segregates 
in the population, effectively changing  the badges expressed and the 
communication system.

```{r model-struc,echo=FALSE,fig.width=8,fig.align='center',fig.height=8,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A). Colour scale in C (applies also for D) indicates the quality of the individual"}

@@

We first present the outcome of simulations where we prevent the 
evolution of the sender code (by setting the mutation rate to $0$), 
and assume that all individuals in the population display either 
an uninformative or informative badge size with respect to their quality. 
These simulations show what type of receiver strategy develops through 
a learning process. Figure \ref{fig:model-struc} (C and D) shows 
the receiver strategy developed through learning; panel C is for 
receivers that faced uninformative signals, and D informative ones. 
When learners face uninformative signals, they change their 
probability of retreating depending on their own quality. Individuals 
with higher quality (red tones), after the learning process, have a 
high probability of attacking; while individuals of lower quality 
(blue tones) mostly retreat from confrontations. Thus, learning splits
the population of receivers into the two classic pure strategies of 
hawks and doves. Given that we have a assumed a monomorphic population 
with unresponsive reaction norms on the signalling side, the changes 
triggered by learning only affect a small range of badge sizes. 
In contrast, when receivers face informative reaction norms 
on the side of the signaller (panel B in fig. \ref{fig:model-struc}), 
receivers use the badge size of their interacting partners to determine 
whether to retreat or attack. The relationship is given by a 
threshold-like reaction norm, where the change from retreating to 
attacking depends on the quality of the receiver. As expected, 
the higher the quality of the receiver the larger the badge size 
that triggers a retreat.

```{r handicap,echo=FALSE,out.width='90%',fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning. Middle panels show the evolutionary dynamics of the sender code. On the left, changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right changes in the distribution of the slope ($\\beta$). Dashed lines in both panels show the mean of the distribution. Grey lines show the generation time corresponding to the panels above and below showing the sender and receiver code. In the panels above, the reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \\ref{fig:model-struc}. In the panels below, colors represent a cluster clasification performed by the k-means algorithm" }

knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn0_betCost5.png"))

@@

When we allow the evolution of the sender code ($\alpha$ and $\beta$ 
change subject to natural selection and genetic drift), and the badge 
size work as a handicap (i.e. the cost of the badge is inversely 
proportional to the quality of the individual), the sender code evolves 
to produce an honest signal Fig \ref{fig:handicap}). The evolution of 
the sender code does not happen immediately after the start of the 
evolutionary process. Instead, the evolutionary dynamics of the reaction 
norm parameters (Fig. \ref{fig:handicap}, X) show a set of 
steps. First, the intercept of the reaction norm ($\alpha$) evolves to 
higher values, reducing the average badge size in the populations. 
That makes sense because reaction norms segregating in the population are 
flat at this point, and consequently the receiver code does not respond to 
the badge size, bigger badges are costlier and do not trigger lower attack 
probabilities. During those first generations, the slope of the sender 
reaction norm ($\beta$)  remains around the initial value of zero. 
At around generation 4000, the slope evolves toward positive values. 
Which implies that larger badges correlate with higher quality. Receivers 
learn to react to such correlation, reducing the probability of attack towards 
individual of larger badges. Hence, natural selection favours larger 
values of $\beta$, eventually leading to an evolutionary equilibrium where 
badge size is an honest signal of quality mediated by the learned responses 
of receivers. The evolutionary trajectory portrayed in figure 
\ref{fig:handicap} is not the only possible trajectory. If the slope of the 
sender reaction norms ($\beta$) evolves toward negative values before badges 
become handicaps, receivers never learn to react to the size of badges. 
Thus, badges are only costly and do not provide information about the 
quality of an individual(Fig. \ref{fig:noBadge} . Eventually, the badge disappears from the population.

When production of the badge is not costly (i.e. does not work as a handicap), 
the parameters of the sender code go through evolutionary branching event, 
triggering the evolution of different types of signals in the population. 
Fig. \ref{fig:branching}A and B shows changes in the distribution of the 
sender reaction norm ($\alpha$ and $\beta$) along evolutionary time. 
Darker areas show trait values with high frequency in the population. 
Populations start monomorphic with a value of zero on both traits, 
and mutations quickly build up a normal distribution around the starting 
value. Within the first 2000 generations the unimodal distribution splits 
into a bimodal one. In the case of $\alpha$, one of the peaks splits further, 
so at the end of the evolutionary simulation the distribution of $\alpha$ 
value in the population shows three peaks. In the case of $\beta$, the two 
initial peaks coexist for around 4000 generations. Later on one of the 
types goes extinct and the distribution goes back to being unimodal. 
These changes in the distribution of the sender reaction norm parameters 
imply that during the second half of the simulation individuals can be 
classified into three distinct types. Two types express a flat reaction 
norm, meaning that their badge size is not informative of their quality, 
while the third type shows intermediate badge sizes which are determined 
by the quality of the individual (Fig. \ref{fig:branching}). Furthermore, 
the receiver reaction norms develop through learning, particularly those 
of individuals of intermediate quality, respond to the signal of their 
sender type by increasing the probability of retreating from a fight with 
individuals with larger badges (Fig. \ref{fig:branching}).

```{r branching,echo=FALSE,out.width='90%',fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The middle panels show changes in the distribution of values fir $\\alpha$ and $\\beta$ along evolutionary time. Panels above and bellow correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}

>>>>

---
title: |
  | Learning can mediate both badges of status and individual recognition
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
- C. Daniel Cadena*
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: 
    latex_engine: xelatex  # word_document: default
fig_caption: yes
fontsize: 12pt
header-includes: |
  ```{=latex}
    \usepackage{float}
    \usepackage{unicode-math}
    \usepackage{setspace}\doublespacing
    \usepackage[backref=page]{hyperref}
    \linespread{2}
    \usepackage{lineno}
    \linenumbers
    \floatplacement{figure}{H}
    \newcommand{\beginsupplement}{ \setcounter{table}{0}         \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  ```
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: ../SocialMorphCue.bib
spacing: double
style: C:\Users\andre\Zotero\styles\elife.csl
abstract: | 
  Adaptive behavioural responses often depend on qualities of the interacting partner of an individual. For example, when competing for resources, an individual might be better off escalating fights with individuals of lower quality, while restraining from fighting individuals of higher quality. Communication systems involving signals of quality allow individuals to reduce uncertainty regarding the fighting ability of their partners and make more adaptive behavioral decisions. However, dishonest individuals can destabilize such communications systems. One open question is whether cognitive mechanisms, such as learning, can maintain the honesty of signals, thus, favoring their evolutionary stability. Here we present evolutionary simulations, where individuals can produce a signal proportional to their quality; and learn along their lifetime the best response to the signal emitted by their peers. In these simulations, learning on the receiver side can mediate the evolution of signals of quality on the sender side. When the cost of the signal is proportional to the quality of the sender populations are only composed of honest signalers. When the cost is not proportional to the quality of the signaler, the population is composed of both honest and dishonest signalers. We argue that learning can be a general cognitive mechanism playing a role in a wide range of communication systems.  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,cache=TRUE,warnings=FALSE)
require(png)
require(here)
library(dplyr)
source(here("../R_files/posPlots.r"))
source(here("AccFunc.r"))
library("scales")
```

@@

The decision of whether to escalate a fight against an interacting 
partner can be dependent on the size of the partner's badge, 
the actual dependency is determined by the focal´s experiences 
through a learning process. We implement learning using 
the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
Individuals estimate the reward (pay-off) expected from  interacting 
with partners of different badge sizes (the critic in RL terminology). 
After every interaction they update the estimate of reward proportionally 
to the difference between their current estimate and the observed 
reward (prediction error $\delta$) and to the speed of learning 
($\mathrm{A}$). Furthermore, they express different probabilities of 
retreating (attacking) depending on the badge size of their opponent 
(the actor in RL). They update the probability of 
retreating (attacking)  up or down depending on whether retreating 
(attacking) leads to an increase in the reward estimation. 
So, if a focal individual decides to escalate a fight against 
an individual with small badge, and that leads to an increase in 
the reward estimation, the focal individual will increase the probability 
of escalating fights with individuals of small badges in the future. 
Given that badge size is a real number between 0 and 1, there are infinitely 
many badge sizes. Thus, the reward estimation, as well as the probability 
of retreating (escalating), must be generalized across different values. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018]. 
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]). This feature centres are the same for all individuals and stored in vector $\symbf{b}$ . Each one of these features centres 
is associated with a weight for reward estimation and tendency to play 
retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights 
associated with each feature, weighted by the response 
triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
represent the feature weights) The response of each feature centre 
diminishes as a Gaussian function  with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:model-struc} B, grey line). Formally the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,
\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

@@

\begin{equation}
logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

@@

After all their interactions, individuals in the population 
reproduce proportionally to their total pay-off $w_i$, which is a sum of the 
baseline pay-off ($w_0$) and all the pay-offs obtained throughout their 
life. Thus, the combination of natural selection and genetic drift 
changes the distribution of values in $\alpha$ and $\beta$ that segregates 
in the population, effectively changing  the badges expressed and the 
communication system.

```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A). Colour scale in C (applies also for D) indicates the quality of the individual"}

@@

We first present the outcome of simulations where we prevent the 
evolution of the sender code (by setting the mutation rate to $0$), 
and assume that all individuals in the population display either 
an uninformative or informative badge size with respect to their quality. 
These simulations show what type of receiver strategy develops through 
a learning process. Figure \ref{fig:model-struc} (C and D) shows 
the receiver strategy developed through learning; panel C is for 
receivers that faced uninformative signals, and D informative ones. 
When learners face uninformative signals, they change their 
probability of retreating depending on their own quality. Individuals 
with higher quality (red tones) after the learning process have a 
high probability of attacking; while individuals of lower quality 
(blue tones) mostly retreat from confrontations. Thus, learning splits
the population of receivers into the two classic pure strategies of 
hawks and doves. Given that we have a assumed a monomorphic population 
with unresponsive reaction norms on the signalling side, the changes 
triggered by learning only affect a small range of badge sizes. 
In contrast, when receivers face informative reaction norms 
on the side of the signaller (panel B in fig. \ref{fig:model-struc}), 
receivers use the badge size of their interacting partners to determine 
whether to retreat or attack. The relationship is given by a 
threshold-like reaction norm, where the change from retreating to 
attacking depends on the quality of the receiver. As expected, 
the higher the quality of the receiver the larger the badge size 
that triggers a retreat.

```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning. Middle panels show the evolutionary dynamics of the sender code. On the left, changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right changes in the distribution of the slope ($\\beta$). Dashed lines in both panels show the mean of the distribution. Grey lines show the generation time corresponding to the panels above and below showing the sender and receiver code. In the panels above, the reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \\ref{fig:model-struc}. In the panels below, colors represent a cluster clasification performed by the k-means algorithm" ,warning=FALSE}

scenario<-"betCostEvol3"#"alphaAct"#"nIntGroupNormQual"#
runChoi<-9 #1

# Load files -------------------------------------------------------------------
# Project folder
listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
evolList<-grep("evolLearn",listTest,value=TRUE)
indList<-grep("indLearn",listTest,value=TRUE)
paramName<-list.files(here("Simulations",paste0(scenario,"_")))
# paramName<-list.files(extSimsDir,full.names = TRUE)
paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*5_*",value = TRUE)

param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
  # fromJSON(paramName[1])
val<-2
fileId<-val
evolList_runs<-grep(paste0(param$namParam,param$rangParam[2]),
                    evolList,value =TRUE)
indList_runs<-grep(paste0(param$namParam,param$rangParam[2]),
                   indList,value =TRUE)
evol<-do.call(rbind,lapply(evolList_runs,fread))
pop<-do.call(rbind,lapply(indList_runs, fread))

Valpar<-param$rangParam[2]

nampar<-param$namParam

nCenters<-param$nCenters
sigSquar<-param$sigSq

# Aesthetic paremeters ---------------------------------------------------------

## Calculate clustering for all generations  -------------------

# # Choose which interaction to visualize
# lastInt<-tail(pop[,unique(nInteract)],4)[2]
# # # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
# # # to get last interaction: tail(pop[,unique(nInteract)],1)
# #
# popOneInd<-pop[nInteract==lastInt]
# vars<-c("alpha","beta","Badge")
# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 5,
#                                 Bsamples =500,iterMax = 500)
# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
#                        by=.(idClust,time,seed)]
# clusSummary[,orderClus:=0]
# invisible(capture.output(
#   for(i in 1:dim(clusSummary)[1]){
#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
#         sort(clusSummary[time==clusSummary[i,time]&
#                            seed==clusSummary[i,seed]]$meanAlph)))
# }
# ))
# 
# 
# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
# 
# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
#                              paste0("popOneInd",nampar,Valpar,".txt")))

popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
paste0("popOneInd",nampar,Valpar,".txt")))

# Get stats from the evolutionary simulations ----------------------------------
# names(evol)

cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
        "meanAlpha", "meanBeta",
        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
        "WeightAct_0","WeightCrit_0",
        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
        "WeightAct_5","WeightCrit_5")

my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
                              upIQR = fivenum(x)[4])

evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
                .SDcols = cols,by=time]

## Plot mean and IQRs among individuals for each replicate --------------------

# get the trajectories for individual runs
traitsTrajs<-dcast(evol,time~seed,
                   value.var = c("meanAlpha","meanBeta",
                                 #"meanInitCrit",#"meanFit",
                                 # "meanInitAct","sdInitCrit","sdInitAct",
                                 "sdAlpha","sdBeta","freqHH",
                                 "freqHD","freqDD"))

finReps<-evol[time==max(time),seed]

genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]

# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
cexAxis<-0.9

nInd2print<-40

par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))

adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
         pal = pal_dist,
         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
         xlim=c(0,22000))

mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
       bty = "n",cex=cexAxis)
legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

# grey lines to show the generations shown in the upper and lower panels

matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
         lty=1,col = "grey",lwd=2)

# Variation among individuals

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
         col=colGenesLin,lty = 2,type="l",lwd=5)

par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    xaxt="s",las=1,new=TRUE)

adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
       col=colGenesLin[2],lwd=2,bty = "n")
legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
# grey lines to show the generations shown in the upper and lower panels
matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,beta]))),
                      lty=1,col = "grey",lwd=2)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

axis(4,cex.axis=cexAxis)

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanBeta_"),runChoi)],
         col=colGenesLin[2],lty = 2,type="l",lwd=4)

# Choose time range
gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
# Plor the actor
seqYax<-c("s",rep("n",3))
# seqYlabUp<-c("Badge",rep("",3))
seqYlabUp<-c("P(dove)",rep("",3))
seqYlabDown<-c("Badge",rep("",3))
seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
seqXlabUp<-c("","Badge","")
labPanels<-c("C","D","E","F","G")
rangQual<-seq(0,1,length.out = 50)
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
rangx<-seq(0,1,length=50)
count<-0
for(genC in genstoPrint){
  count<-count+1
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
                  lowboundx = 8, title = TRUE,lowboundy = 10,
                  upboundy = 87),
      xaxt="s",las=1,new=TRUE)
  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
                             .SDcols=grep("WeightAct",
                                          names(evol),value = TRUE)])
  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
               .SDcol=c(grep("WeightAct",
                             names(evol),value = TRUE),"Quality","alpha","beta",
                        "orderClus"),
               by=indId]

  randInds<-sample(tempPop$indId, nInd2print, replace = FALSE)

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],
                         colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("D.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
#          beta = meansClustmp[orderClus==x,betaMean])})
# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
#          type='l',cex.axis=cexAxis,
#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
    logist(rangx,alpha = tempPop[indId==x,alpha],
           beta = tempPop[indId==x,beta])})
  matplot(x = rangx,y = reacNormInd,
          type='l',cex.axis=cexAxis,
          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
  legend("topleft",legend =  c(paste0("C.",count)),bty = "n",inset = -0.06)
  }

rm(list=grep("temp",ls(),value = T))

@@

When we allow the evolution of the sender code ($\alpha$ and $\beta$ 
change subject to natural selection and genetic drift), and the badge 
size work as a handicap (i.e. the cost of the badge is inversely 
proportional to the quality of the individual), the sender code can evolve 
to produce an honest signal Fig \ref{fig:handicap}). The evolution of 
the sender code does not happen immediately after the start of the 
evolutionary process. Instead, the evolutionary dynamics of the reaction 
norm parameters (Fig. \ref{fig:handicap}, X) show a set of 
steps. First, the intercept of the reaction norm ($\alpha$) evolves to 
higher values, reducing the average badge size in the populations. 
That makes sense because reaction norms segregating in the population are 
flat at this point, and consequently the receiver code does not respond to 
the badge size, bigger badges are costlier and do not trigger lower attack 
probabilities. During those first generations, the slope of the sender 
reaction norm ($\beta$)  remains around the initial value of zero. 
At around generation 4000, the slope evolves toward positive values. 
Which implies that larger badges correlate with higher quality. Receivers 
learn to react to such correlation, reducing the probability of attack towards 
individual of larger badges. Hence, natural selection favours larger 
values of $\beta$, eventually leading to an evolutionary equilibrium where 
badge size is an honest signal of quality mediated by the learned responses 
of receivers. The evolutionary trajectory portrayed in figure 
\ref{fig:handicap} is not the only possible trajectory. If the slope of the 
sender reaction norms ($\beta$) evolves toward negative values before badges 
become handicaps, receivers never learn to react to the size of badges. 
Thus, badges are only costly and do not provide information about the 
quality of an individual (Fig. \ref{fig:noBadge} . Eventually, the badge 
disappears from the population. In contrast, when the badge does not work as a handicap but instead is cost-free, the evolutionary process never leads to the 
establishment of an honest signal. Instead, subject to genetic drift the 
population either evolves towards the disappearance of the badge, or to the 
maximum badge size. In either of this cases reaction norms are flat, so the 
badge does not provide any quality information.

The amount of information that agents are able to collect through learning 
throughout their life time can strongly change the outcome of 
the evolutionary dynamics. In the simulations presented in figure 
\ref{fig:handicap} and \ref{fig:noBadge}, 
individuals learned with high learning rate ($\alpha = 0.4$) and interacted 
repeatedly along their lifetime (on average 2000 interactions). When we reduce 
the number of interactions that individuals have over their lifetime to 300 
interactions on average, we see an drastic increase in the morphological and genetic
variation present in the population. Fig. \ref{fig:branching} shows 
changes in the distribution of the sender reaction norm
along evolutionary time assuming the badge is cost-free. 
Darker areas show trait values with high frequency in the population. 
Populations start monomorphic with a value of zero on both traits, 
and mutations quickly build up a normal distribution around the starting 
value. Within the first 2000 generations the unimodal distribution in 
the intercept ($\alpha$) splits into a bimodal one. 
Later in evolutionary time one of the peaks splits further, 
so at the end of the evolutionary simulation the distribution of $\alpha$ 
value in the population shows three peaks. In the case of $\beta$, the 
peaks in the distribution are not so clear-cut, but it's clear that there is an 
increase in the variance of the distribution. These changes in the 
distribution of the sender reaction norm parameters 
imply that during the second half of the simulation individuals can be 
classified into three distinct types. Two types express a flat reaction 
norm, meaning that their badge size is not informative of their quality, 
while the third type shows intermediate badge sizes which are determined 
by the quality of the individual (Fig. \ref{fig:branching}). Furthermore, 
the receiver reaction norms develop through learning, particularly those 
of individuals of intermediate quality, respond to the signal of their 
sender type by increasing the probability of retreating from a fight with 
individuals with larger badges (Fig. \ref{fig:branching}).
